LME-MPM applied to quasi-brittle fracture.

### 2020-05-25 2020_EFM_MPM_Eigensoftening.tex
Cambios 25-5-2020
Nueva sección describiendo el algoritmo de reconstrucción de la vecindad. 
Aclaraciones sobre la notación

@@

\subsection{The \acrshort{mpm} time integration : A \acrlong{npc}  scheme}
\label{sec:2.1}
The \acrshort{mpm}~\cite{Sulsky1994} is a meshfree Lagrangian-Eulerian
method where moving material points, henceforth in this research, particles will carry on
all the physical information of the local state ($\sigma_p,
\varepsilon_p$) and a set of fixed background nodes will be introduced to compute the balance of momentum equation. Since the \acrshort{mpm} 
possesses the advantages of both Lagrangian and Eulerian descriptions,
no element distortion takes place in the \acrshort{mpm}. Therefore, it
is an appropriate and efficient method to solve problems with moving
discontinuities such as fracture evolution.\\
Without loosing generality, the \acrshort{mpm} algorithm
can be described with three main steps: (i) a variational recovery
process, where particle data is projected to the grid nodes, (ii) an
Eulerian step, where balance of momentum equation is expressed as a
nodal equilibrium equation in a FEM-like procedure, and finally
(iii) a Lagrangian advection of the particles. In the present research,
an explicit predictor-corrector time integration scheme is adopted. The
purposes of this choice is motivated due to its proved robustness and
stability for dynamic computations. In the first stage, the nodal
velocity predictor is computed following \eqref{eq:Predictor-velocity}: 
\begin{equation}
  \label{eq:Predictor-velocity}
  \vec{\tilde{v}}_I^{k+1} = \frac{ N_{Ip}^{k} m_p (\vec{v}_p^k + (1 - \gamma)\ \Delta t\ \vec{a}_p^k)}{\tens{m}_I^{k+1}}\ .
\end{equation}
\PNA{EXPLICAR NOTACION, CUIDADO CON EL GAMMA} This way of computing the nodal predictor is numerically stable
as well as minimises the computational effort. Once nodal velocities are
obtained, the essential boundary conditions are imposed. After that, an
Eulerian phase is computed in the set of nodes in a
\acrshort{fem}-like way, where nodal forces $\vec{f}_{I}^{k+1}$ are
computed through the equilibrium equation. Next the nodal velocities
are corrected in a \textit{corrector} stage:
\begin{equation}
  \label{eq:Corrector-velocity}
  \vec{v}_{I}^{k+1} = \vec{v}_{I}^{pred} + \gamma\ \Delta t\
  \frac{\vec{f}_{I}^{k+1}}{\tens{m}_I^{k+1}}\ .
\end{equation}
Finally, in order to update the particles, those are advected in the Lagrangian stage as
\begin{align}
  \label{eq:Update-lagrangian-pce}
        &\vec{a}_p^{k+1} =
          \frac{N_{Ip}^k\vec{f}_{I}^{k}}{\tens{m}_I^k}\ ,\\
      &\vec{v}_p^{k+1} = \vec{v}_p^n + \Delta t\
        \frac{N_{Ip}^k\
        \vec{f}_{I}^{k}}{\tens{m}_I^k}\ ,\\
      &\vec{x}_p^{k+1} = \vec{x}_p^n + \Delta t\
         N_{Ip}^k\ \vec{v}_{I}^{k} +
        \frac{1}{2}\Delta t^2\ \frac{N_{Ip}^k\
        \vec{f}_{I}^{k}}{\tens{m}_I^k}\ .
\end{align}
The complete pseudo-algorithm it is summarised in \ref{sec:expl-pred-corr}.

@@

(0,\infty)$. Notice that $\beta$ has units of $[L]^{-2}$,
therefore it can be controlled by adjusting a dimensionless parameter,
$\gamma=\beta h^2$ \cite{Arroyo2006}, where $h$ is defined as a
suitable measure of the nodal spacing. With the restrictions exposed in
\eqref{eq:LME-scheme-pareto-set}, the unique solution of the local

@@

Implementation details can be consulted in \ref{sec:eigens-algor-1}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Cases of study and discussion}
\label{sec:3}

@@

\begin{table}
  \centering
  \begin{tabular}[]{c c c c c c c}
    \hline
      &   $\rho$   & $\text{f}_c$ & $\text{f}_t$ & $G_F$ &   E   & $\text{d}_{max}$ \\
      & (kg/m$^3$) &     (MPa)    &     (MPa)    & (N/m) & (GPa) & (mm) \\
    \hline
Value &    2368    &     102.7    &      5.4     &  141  &   31  &  12 \\
    \hline
  \end{tabular}
  \caption[Mechanical properties of thje concrete]{Mechanical
    properties of the high strenght concrete.}
  \label{tab:concrete-properties}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
For the present research, a 2D setting of the experiment has been
adopted. Both the hammer and the concrete beam are explicitly
represented. Several levels of discretization were required to assess
the objectiveness of the obtained results. The results here exposed
are from a discretization of 30616 material points and 6297 nodes. To
achieve better results, an unestructured grid layout has been adopted
focusing the minimum nodal spacing, 0.47 mm, in the middle of the beam and in
the impact surface of the hammer. For this simulation, $\gamma$ was
fixed to 6.\\

@@

\begin{algorithm}[H]
  \DontPrintSemicolon
  \KwInput{For each $p$, $\epsilon$-neighbourhood, $f_{t,p}$, $h_{\epsilon,p}$, $w_c$}
  \KwOutput{Return damage parameter $\chi := \{\chi_p\}$}
  % \KwData{Testing set $x$}
  $\chi_p \leftarrow \chi_p^{k}$ \;
  \For{$p$ to $N_p$}
  {
    \If{$\chi_p = 0$ \AND $\epsilon_{f,p} = 0$}
    {
      \For{$q \in B_{\epsilon,p}$}
      {
        $\sigma_{q,I} \leftarrow \text{getEigenvaluesOf}(\sigma_{q})$ \;
        \If{$\chi_q < 1$}
        {
          $\sum m_p\sigma_{p,I} \leftarrow \sum m_p\sigma_{p,I} + m_q\sigma_{q,I}$ \;
        }
        $m_p \leftarrow m_p + m_q$ \;
      }
      $\sigma_{p,\epsilon} \leftarrow \frac{1}{m_p} \sum m_p\sigma_{p,I}$\;
      \If{$\sigma_{p,\epsilon} > f_{t,p}$}
      {
        $\epsilon_{f,p} = \epsilon_{I,p}$ \;  
      }
      \ElseIf{$\chi_p \neq 1$ \AND $\epsilon_{f,p} > 0$}
      {
        $\chi_p^{k+1} \leftarrow \min\Big \{1 , \max \{\chi_p^{k},
        \frac{(\epsilon_{I,p}- \epsilon_{f,p})\ h_{\epsilon,p}}{w_c}
        \} \Big \}$ \;
      }
    }
  }
  \caption{Compute damage parameter $\chi_p^{k+1}$}
  \label{alg-eigens}
\end{algorithm}

>>>>

\newglossaryentry{gamma-npc}
{
  name=$\gamma$,
  description={Time integration parameter for the \acrshort{npc} algorithm}
}

\newglossaryentry{gamma-lme}
{
  name=$\widehat{\gamma}$,
  description={Dimensionless parameter to control the value of the
    thermalization parameter $\beta$ of the \acrshort{lme} shape functions.}
}

@@

\subsection{The \acrshort{mpm} time integration : A \acrlong{npc}  scheme}
\label{sec:2.1}
The \acrshort{mpm}~\cite{Sulsky1994} is a meshfree Lagrangian-Eulerian
method where moving material points, henceforth in this research,
particles \MMP{($\square_p$)} will carry on all the physical information of the local state ($\sigma_p,
\varepsilon_p$) and a set of fixed background nodes \MMP{($\square_I$)} will
be introduced to compute the balance of momentum equation. Since the
\acrshort{mpm} possesses the advantages of both Lagrangian and
Eulerian descriptions, no element distortion takes place in the
\acrshort{mpm}. Therefore, it is an appropriate and efficient method
to solve problems with moving discontinuities such as fracture evolution.\\
Without loosing generality, the \acrshort{mpm} algorithm
can be described with three main steps: (i) a variational recovery
process, where particle data is projected to the grid nodes, (ii) an
Eulerian step, where balance of momentum equation is expressed as a
nodal equilibrium equation in a FEM-like procedure, and finally
(iii) a Lagrangian advection of the particles. In the present research,
an explicit predictor-corrector time integration scheme is adopted. The
purposes of this choice is motivated due to its proved robustness and
stability for dynamic computations. \MMP{In the first stage, the nodal
velocity predictor $ \vec{\tilde{v}}_I^{k+1}$ is computed thorough
nodal velocities $\vec{v}_I$ and accelerations $\vec{a}_I$ evaluated
in the $k$ time step},
\begin{equation}
  \label{eq:Predictor-velocity}
  \vec{\tilde{v}}_I^{k+1} = \frac{N_{Ip}^{k} m_p (\vec{v}_p^k + (1 - \gamma)\ \Delta t\ \vec{a}_p^k)}{\tens{m}_I^{k+1}}\ .
\end{equation}
\MMP{Where \gls{gamma-npc} is an user-defined parameter (typically adopted as
0.5), $\tens{m}_I$ represents the lumped mass matrix, and $N_{Ip}^{k}$ is
the nodal contribution of the shape function evaluated in the particle
coordinates $\vec{x}_p^k$.} This way of computing the nodal predictor is numerically stable
as well as minimises the computational effort. Once nodal velocities are
obtained, the essential boundary conditions are imposed. After that, an
Eulerian phase is computed in the set of nodes in a
\acrshort{fem}-like way, where nodal forces $\vec{f}_{I}^{k+1}$ are
computed through the equilibrium equation. Next the nodal velocities
are corrected in a \textit{corrector} stage:
\begin{equation}
  \label{eq:Corrector-velocity}
  \vec{v}_{I}^{k+1} = \vec{v}_{I}^{pred} + \gamma\ \Delta t\
  \frac{\vec{f}_{I}^{k+1}}{\tens{m}_I^{k+1}}\ .
\end{equation}
Finally, in order to update the particles, those are advected in the Lagrangian stage as
\begin{align}
  \label{eq:Update-lagrangian-pce}
        &\vec{a}_p^{k+1} =
          \frac{N_{Ip}^k\vec{f}_{I}^{k}}{\tens{m}_I^k}\ ,\\
      &\vec{v}_p^{k+1} = \vec{v}_p^n + \Delta t\
        \frac{N_{Ip}^k\
        \vec{f}_{I}^{k}}{\tens{m}_I^k}\ ,\\
      &\vec{x}_p^{k+1} = \vec{x}_p^n + \Delta t\
         N_{Ip}^k\ \vec{v}_{I}^{k} +
        \frac{1}{2}\Delta t^2\ \frac{N_{Ip}^k\
        \vec{f}_{I}^{k}}{\tens{m}_I^k}\ .
\end{align}
The complete pseudo-algorithm it is summarised in \ref{sec:expl-pred-corr}.

@@

(0,\infty)$.\MMP{Notice that $\beta$ has units of $[L]^{-2}$,
therefore it can be controlled by adjusting a dimensionless
parameter\footnote{To avoid confusion with the \gls{gamma-npc} parameter of
  the \acrshort{npc}, the dimensionless parameter defined in Arroyo \&
  Ortiz \cite{Arroyo2006} as $\gamma$ will be represent by
  \gls{gamma-lme} to preserve as much as possible the original notation.},
$\widehat{\gamma}=\beta h^2$}, where $h$ is defined as a
suitable measure of the nodal spacing. With the restrictions of the
Pareto set, the unique solution of the local

@@

Implementation details can be consulted in \ref{sec:eigens-algor-1}.

\MMP{
\subsection{$\epsilon$-neighborhood reconstruction : A node-linked method}
\label{sec:epsil-neighb-reconst}
Since this operation could be extremely demanding, optimal
numerical implementation should be employed. In opposite with the Cell-linked
method proposed by Allen \& Tildesley \cite{Allen_et_al_1989} where
the definition of a numerical tolerance is required for auxiliar
algorithms like Crossing Number \cite{Shimrat_1962} or Correct
Even-Odd \cite{Galetzka_et_al_2017}. To avoid these error prone
techniques, the method here proposed exploits the meshfree benefits of
the \acrshort{lme} approximants to introduce a node-linked method. Due
to the global definition of this interpolation technique, the classical
interpretation of the \acrshort{fem} element typically adopted in
\acrshort{mpm} simulations is not required anymore, instead a linked
list of tributary nodes is reconstructed with the closest node as
starting point at each time step. Therefore, each node is assigned an
unique linked-list of those particles which are closest to it. This
approach allows to reconstruct easily a search list with the support
of the initial connectivity of the mesh and offers a suitable basis
for an easy implementation of other interpolation techniques with
minimum coding effort. A detailed explanation of the proposed
algorithm can be found in \ref{sec:node-linked-method-1}.
}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Cases of study and discussion}
\label{sec:3}

@@

\begin{table}
  \centering
  \begin{tabular}[]{c c c c c c c}
    \hline
      &   $\rho$   & $\text{f}_c$ & $\text{f}_t$ & $G_F$ &   E   & $\text{d}_{max}$ \\
      & (kg/m$^3$) &     (MPa)    &     (MPa)    & (N/m) & (GPa) & (mm) \\
    \hline
Value &    2368    &     102.7    &      5.4     &  141  &   31  &  12 \\
    \hline
  \end{tabular}
  \caption[Mechanical properties of thje concrete]{Mechanical
    properties of the high strenght concrete.}
  \label{tab:concrete-properties}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
For the present research, a 2D setting of the experiment has been
adopted. Both the hammer and the concrete beam are explicitly
represented. Several levels of discretization were required to assess
the objectiveness of the obtained results. The results here exposed
are from a discretization of 30616 material points and 6297 nodes. To
achieve better results, an unestructured grid layout has been adopted
focusing the minimum nodal spacing, 0.47 mm, in the middle of the beam and in
the impact surface of the hammer. For this simulation, \gls{gamma-lme} was
fixed to 6.\\

@@

\begin{algorithm}[H]
  \DontPrintSemicolon
  \KwInput{For each $p$, $\epsilon$-neighbourhood, $f_{t,p}$, $h_{\epsilon,p}$, $w_c$}
  \KwOutput{Return damage parameter $\chi := \{\chi_p\}$}
  % \KwData{Testing set $x$}
  $\chi_p \leftarrow \chi_p^{k}$ \;
  \For{$p$ to $N_p$}
  {
    \If{$\chi_p = 0$ \AND $\epsilon_{f,p} = 0$}
    {
      \For{$q \in B_{\epsilon,p}$}
      {
        $\sigma_{q,I} \leftarrow \text{getEigenvaluesOf}(\sigma_{q})$ \;
        \If{$\chi_q < 1$}
        {
          $\sum m_p\sigma_{p,I} \leftarrow \sum m_p\sigma_{p,I} + m_q\sigma_{q,I}$ \;
        }
        $m_p \leftarrow m_p + m_q$ \;
      }
      $\sigma_{p,\epsilon} \leftarrow \frac{1}{m_p} \sum m_p\sigma_{p,I}$\;
      \If{$\sigma_{p,\epsilon} > f_{t,p}$}
      {
        $\epsilon_{f,p} = \epsilon_{I,p}$ \;  
      }
      \ElseIf{$\chi_p \neq 1$ \AND $\epsilon_{f,p} > 0$}
      {
        $\chi_p^{k+1} \leftarrow \min\Big \{1 , \max \{\chi_p^{k},
        \frac{(\epsilon_{I,p}- \epsilon_{f,p})\ h_{\epsilon,p}}{w_c}
        \} \Big \}$ \;
      }
    }
  }
  \caption{Compute damage parameter $\chi_p^{k+1}$}
  \label{alg-eigens}
  \caption{Eigensoftening algorithm}
\end{algorithm}

\section{Node-linked method for $\epsilon$-neighborhood reconstruction}
\label{sec:node-linked-method-1}
\begin{algorithm}[H]
  \DontPrintSemicolon
  \KwInput{$I_0$ : Closest node for each particle}
  \KwInput{$I_{Nodes}$ : List of nodes close to each node}
  \KwInput{$Particles_I$ : List of particles close to each node}
  \KwOutput{$\text{neighborhood}_{\epsilon}$ : $\epsilon$-neighborhood for each particle}
  \For{$p$ to $N_p$}
  {
    For each $p$ get $I_0$\;
    Eval $I_{Nodes}(I_0)$ \;
    \For{$k \in I_{Nodes}(I_0)$}
    {
      $\text{neighborhood}_{\epsilon}^{k+1}$ = $\text{neighborhood}_{\epsilon}^{k} \cup Particles_I(k)$
    }
    \For{$p_{\epsilon} \in \text{neighborhood}_{\epsilon}$}
    {
      \If{$distance(p_{\epsilon},p) > \epsilon $}
      {
        pop $p_{\epsilon}$ from neighborhood$_{\epsilon}$\;
      }      
    }

  }  
  \caption{$\epsilon$-neighborhood reconstruction algorithm}
\end{algorithm}
